Data Exploration: IUCN and AquaMaps

### 2016-08-07 coralmaps/coral_maps.Rmd
cleaning up the old drafts; cleaning up the data_explores

  coral_spp_sid <- spp_map_pairs %>%
    filter(str_detect(spp_group, 'CORAL')) %>%
    .$iucn_sid

@@

  spp_depth <- spp_narr %>%
    select(iucn_sid = species_id, habitat)
  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))
  ### manually find max depths from this
}

@@

spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>%
  mutate(deep = 1) %>%
  select(-weight)

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
iucn_coral_cells <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  select(iucn_sid, loiczid = LOICZID) %>%
  filter(iucn_sid %in% spp_depth$iucn_sid) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

write_csv(iucn_coral_cells, file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))

``` {r}
### calc areas
### read hcaf to get areas; left join to iucn_coral_cells
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
iucn_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))

@@

iucn_coral_areas <- iucn_coral_cells %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(deeparea = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(deeparea),
         shallow_pct = (totarea - deeparea)/totarea) %>%
  filter(deep == 1) %>%
  select(-deep)

@@

``` {r}
if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))) {

  ### get species list
  iucn_coral_cells_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% iucn_coral_cells_clipped$iucn_sid) %>%
    select(am_sid, sciname, iucn_sid)

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))) {
    am_coral_cells <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(am_coral_cells, file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
  }

  am_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))

@@

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, iucn_coral_cells_clipped) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, am_coral_cells) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

@@

  print(head(spp_coralmaps))
  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))
}
```

``` {r}
### combine old and new area dataframes
spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv')) %>%
  select(iucn_sid, am_sid, sciname, area_am, 
         area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
         sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
  left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
              select(am_sid, iucn_sid, 
                     area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                     sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
            by = c('iucn_sid', 'am_sid'))

write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
```

@@

clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  geom_line(color = 'grey50', alpha = .5) +
  geom_point(aes(color = method)) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)',
       title = 'Coral species alignment, IUCN full range vs depth-clipped')

clipped_quads
```

>>>>

  ### get paired species list
  paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
  spp_map_pairs <- read_csv(paired_maps_file) %>%
    filter(!is.na(popn_category)) %>% ### filter out unassessed species
    filter(!is.na(sm_perc))           ### filter out NA maps

  ### get full species list
  spp_all_file <- file.path(dir_data, 'spp_all.csv')
  spp_all <- read_csv(spp_all_file)

  coral_spp_sid <- spp_all %>%
    filter(str_detect(tolower(spp_group), 'coral')) %>%
    .$iucn_sid %>%
    unique()

@@

  spp_depth <- spp_narr %>%
    select(iucn_sid = species_id, habitat) %>%
    mutate(paired_map = (iucn_sid %in% spp_map_pairs$iucn_sid))

spp_depth1 <- spp_depth %>%
  mutate(depth = str_extract_all(habitat, '\\d++(?:\\.\\d++)?(?=\\s*+(?:m))')) %>%
  unnest(depth) %>%
  group_by(iucn_sid) %>%
  summarize(max_depth = max(as.integer(depth)))

spp_depth <- spp_depth %>%
  left_join(spp_depth1, by = 'iucn_sid') %>%
  mutate(shallow_str = str_detect(tolower(habitat), 'shallow'),
         shallow = max_depth <= 200 | shallow_str)

write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

summary(spp_depth)
 #    iucn_sid        habitat          paired_map        max_depth      shallow_str     shallow       
 # Min.   :132859   Length:562         Mode :logical   Min.   :  1.00   Mode :logical   Mode:logical  
 # 1st Qu.:133081   Class :character   FALSE:174       1st Qu.: 20.00   FALSE:194       TRUE:544      
 # Median :133284   Mode  :character   TRUE :388       Median : 20.00   TRUE :367       NA's:18       
 # Mean   :133283                      NA's :0         Mean   : 27.55   NA's :1                       
 # 3rd Qu.:133486                                      3rd Qu.: 30.00                                 
 # Max.   :133703                                      Max.   :200.00                                 
 #                                                     NA's   :46

nrow(spp_depth %>% filter(max_depth > 100))
# [1] 5
### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m

spp_depth_na <- spp_depth %>%
  filter(is.na(shallow))

### 17 species (of 396) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
```

``` {r extract_IUCN_cells_for_coral_pairs}

bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>%
  mutate(deep = 1) %>%
  select(-weight)

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
coral_cells_iucn <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  select(iucn_sid, loiczid = LOICZID) %>%
  filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

@@

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

@@

iucn_coral_areas <- coral_cells_iucn %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(deeparea = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(deeparea),
         shallow_pct = (totarea - deeparea)/totarea) %>%
  arrange(desc(deep)) %>%
  .[!duplicated(.$iucn_sid), ]

@@

``` {r calc_overlaps}
if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))) {

  ### get species list
  coral_cells_iucn_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% coral_cells_iucn_clipped$iucn_sid) %>%
    select(am_sid, sciname, iucn_sid)

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))) {
    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(coral_cells_am, file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
  }

  coral_cells_am <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))

@@

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, coral_cells_iucn_clipped) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

@@

  print(head(spp_coralmaps))

  ### combine old and new area dataframes
  spp_coralmaps <- spp_coralmaps %>%
    select(iucn_sid, am_sid, sciname, area_am, 
           area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
           sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
    left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
                select(am_sid, iucn_sid, 
                       area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                       sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
              by = c('iucn_sid', 'am_sid'))

  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
}
```

@@

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#4dac26")  + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#b8e186") + 
  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#f1b6da") + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#d01c8b") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +

  geom_line(color = 'grey50', size = .5, alpha = .3) +
  geom_point(aes(color = method), size = 2, alpha = .8) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)')

print(clipped_quads)

ggsave(plot = clipped_quads,
       filename = file.path(dir_fig, 'scatter_quadmap_corals.png'), device = 'png',
       width = 17.8, height = 9, units = 'cm', dpi = 600)

### 2016-08-07 data_explore/data_explore.Rmd
cleaning up the old drafts; cleaning up the data_explores

iucn_file <- file.path(dir_data, 'spp_iucn.csv')
spp_iucn <- read.csv(iucn_file, stringsAsFactors = FALSE)

message('Getting IUCN maps list')
iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp

@@

sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)

@@

taxa_am_all_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn$sciname) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_ass  = n(),
            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))

taxa_am_solo_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn$sciname) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo_ass = n(),
            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))

taxa_df <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_am_all_assessed,  by = 'class') %>%
  full_join(taxa_am_solo_assessed, by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))

@@

# taxa_df_long1 <- taxa_df %>%
#   dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
#   gather(dataset, n_spp, AquaMaps:Both)
# bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
#   ggtheme_plot + 
#   geom_bar(stat = 'identity') +
#   facet_grid(dataset ~ ., scales = 'free_y') +
#   theme(axis.text.x = element_text(angle = 75, hjust = 1))
# print(bar_taxa1)

@@

taxa_df_long2 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100,
         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
  arrange(desc(dataset))

bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = .8) +
  scale_fill_manual(values = colvec) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = c(0, 25, 50, 75, 100),
                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

print(bar_taxa2)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm', dpi = 600)

# taxa_df_long3 <- taxa_df %>%
#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
#   gather(dataset, n_spp, AquaMaps:Both) %>%
#   group_by(class) %>%
#   mutate(n_total = sum(n_spp, na.rm = TRUE),
#          percent = n_spp/n_total * 100)
#
# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
#   ggtheme_plot + 
#   geom_bar(stat = 'identity', alpha = .5) +
#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
#             #size = 3, 
#             angle = 90, hjust = 0, color = 'grey30') +
#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
#        x = 'Class', 
#        y = 'Percent of assessed mapped species')
# print(bar_taxa3)
# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.tif'), device = 'tiff', 
#        width = 17.8, height = 6, units = 'cm', dpi = 600)

@@

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc)) + #, 
 #                               size = lg_area_pct,
 #                              color = popn_category)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
             fill= "#d01c8b") + 
  geom_point(color = '#4d4dac', alpha = .6)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
#   scale_colour_manual(values = c('LC' = 'green4', 
#                                  'NT' = 'greenyellow', 
#                                  'VU' = 'yellow', 
#                                  'EN' = 'orange', 
#                                  'CR' = 'orangered2',
#                                  'DD' = 'grey30')) + 
#   scale_size_continuous(limits = c(0, 100),
#                         breaks = c( 0,    25,    50,    75,    100), 
#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)')) #, 
#       title = 'Distribution alignment vs Extent alignment',
#       size = '% ocean range',
 #      color = 'Extinction\nRisk')

@@

ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
       width = 17.8, height = 9, units = 'cm', dpi = 600)

@@

spp_gp_quadrants <- spp_gp_quadrants %>%
  left_join(spp_gp_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(spp_group_text, pct_q1 = pct_quad),
            by = 'spp_group_text') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched'))

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .8) +
  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
                               'q2 dist-matched'   = '#b8e186', 
                               'q3 area-matched'   = '#f1b6da', 
                               'q4 poorly matched' = '#d01c8b')) +
#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
#                                'q2 dist-matched'   = 'greenyellow', 
#                                'q3 area-matched'   = 'orange', 
#                                'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = c(0, .25, .5, .75, 1.0),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of species by quadrant', 
#       title = 'Taxonomic group by quadrant',
       fill = 'Alignment')

@@

# Obsolete plots; not evaluated

* bar chart: number of spp per group
* scatter plot: area compare IUCN vs AM, including log plot and plot by species group
* rangemap: Chromis alpha

``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
```

* percent overlap vs total area exploration

``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
```

* Multiple species maps per quadrant

``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
```

>>>>

iucn_file <- file.path(dir_data, 'spp_iucn.csv')
spp_iucn <- read_csv(iucn_file)

message('Getting IUCN maps list')
iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
spp_iucn_maps <- read_csv(iucn_maps_file) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp

@@

sciname_lookup <- read_csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'))

@@

taxa_df <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))

@@

@@

taxa_df_long <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100,
         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
  arrange(desc(dataset))

bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = .9) +
  scale_fill_manual(values = colvec) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = c(0, 25, 50, 75, 100),
                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

print(bar_taxa)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm', dpi = 600)

@@

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc)) + #, 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b")

  ### plot the data points!
scatter_quadplot <- scatter_quadplot +
  geom_point(color = '#4d4dac', alpha = .6)

  ### label the quadrants:
scatter_quadplot <- scatter_quadplot +
  annotate("rect", xmin = 82, xmax = 98, ymin = 85, ymax = 91, 
           alpha = .7, fill= 'white')  + 
  annotate("rect", xmax = 24, xmin =  2, ymin = 85, ymax = 91, 
           alpha = .7, fill= "white") + 
  annotate("rect", xmin = 82, xmax = 98, ymax =  9, ymin = 15, 
           alpha = .7, fill= "white") + 
  annotate("rect", xmax = 20, xmin =  2, ymax =  9, ymin = 15, 
           alpha = .7, fill= "white") +
  annotate("text", x = 96, y = 88, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
           fontface = 'bold.italic', label = "Well-aligned") + 
  annotate("text", x =  4, y = 88, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Distribution-aligned") + 
  annotate("text", x = 96, y = 12, hjust = 1, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Area-aligned") + 
  annotate("text", x =  4, y = 12, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Poorly aligned")

### Manage axis scales 
scatter_quadplot <- scatter_quadplot +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

@@

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'))

@@

ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
       width = 17.8, height = 9, units = 'cm', dpi = 600)

@@

quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
                         'quad_name' = factor(c('well-matched', 'dist-matched', 
                                               'area-matched', 'poorly matched'),
                                             ordered = TRUE))

spp_gp_quadrants <- spp_gp_quadrants %>%
  left_join(spp_gp_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(spp_group_text, pct_q1 = pct_quad),
            by = 'spp_group_text') %>%
  left_join(quad_names, by = 'quad')

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .8) +
  scale_fill_manual(values = c('q1'   = '#4dac26', 
                               'q2'   = '#b8e186', 
                               'q3'   = '#f1b6da', 
                               'q4' = '#d01c8b'),
                    labels = levels(quad_names)) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = c(0, .25, .5, .75, 1.0),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of species by quadrant', 
       fill = 'Alignment')

@@
